---
link-citations: yes
output:
  html_fragment:
    self_contained: false
    includes:
    number_sections: no
---

<!--yaml
---
mermaid: false
mathjax2: false
mathjax: false
highlight: true
title: "Making Jekyll Blog Suitable for R-bloggers"
description: "A demo for fixing problems that may be encountered when one wants to add blog to R-bloggers."
tags: [R-bloggers, Jekyll, R Markdown, Web Page]
---
yaml-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.dim = c(7, 4),
	message = FALSE,
	warning = FALSE,
	comment = "",
	out.width = "100%",
	fig.path = "../../assets/rblogger-criteria-img/"
)
library(knitr)
```

According to the post [add your blog](https://www.r-bloggers.com/add-your-blog/page/5/), adding one's blog to [R-bloggers](https://www.r-bloggers.com/) isn't easy at all, especially for people who use [R Markdown](https://rmarkdown.rstudio.com) to write posts and use [Jekyll](https://jekyllrb.com/) to generate static web page on GitHub.
<!--more-->

Two reasons make it difficult:

1. > The **feed** you submit **should ONLY be about R** (e.g: with R code, or directly related to the R world/community).
    - This causes problem because most bloggers write posts spanning several topics, and some blogging platform doesn't support RSS feed for a particular category of posts.

2. > Make sure the HTML of your feed’s content is well formatted – otherwise it would not be able to load on r-bloggers. This includes to NOT copy-pasting from RStudio’s notebook extension – the feed **should NOT include “base64” images**, make sure your images are saved as png or similar file formats.
    - For posts written with [R Markdown](https://rmarkdown.rstudio.com) and [HTML Fragment](https://rmarkdown.rstudio.com/html_fragment_format.html) output format, the embedded images generated from code chunks (such as outputs from `plot()`) are **base64 images**.


## Fixing the First Problem

For **Jekyll** sites using [jekyll-feed](https://github.com/jekyll/jekyll-feed) plugin to generate sitemaps, it is not possible to have RSS feed for a particular tag or category of posts. However there's a workaround using **Jekyll**'s [Liquid syntax](https://shopify.github.io/liquid/) to write an RSS template, as indicated by [this post](https://devblog.dymel.pl/2017/02/09/category-rss-feed-in-jekyll/). I modified the template to make it suitable for the criteria set by R-bloggers, you can take a look at the file [feed.rbloggers.xml](https://github.com/liao961120/liao961120.github.io/blob/master/feed.rbloggers.xml).


## Fixing the Second Problem

To fix the second problem, set `self_contained: false` in the yaml header of the R Markdown document. With this setting, plots from code chunks are automatically generated in a figure directory, and the `.html` output uses `<img>` tags to source the plots in that figure directory.

```yaml
---
output:
  html_fragment:
    self_contained: false
---
```

However, this creates a problem since the relative path in the `<img>` tags probably won't work in the remote directory that hosts your site. To overcome this problem, you have to **change the default figure directory** and **post-process the output `.html` file**.

For example, the source of this post(`rblogger-criteria.rmd`) is two layers under the root dir of the web site.

```
+---assets/
|   +---rblogger-criteria-img/
+---_includes/
+---_plugins/
+---_posts/
+---post_source/
|   +---rblogger-criteria/
|   |   +---rblogger-criteria.rmd
|   |   +---rblogger-criteria.html
+---index.html
+---_config.yml
```

I set my figure directory to `assets/rblogger-criteria-img/`:

````r
`r ''````{r setup, include=FALSE}
knitr::opts_chunk$set(
	fig.path = "../../assets/rblogger-criteria-img/"
)
```
````

so when the site is rendered, the image would be in `https://liao961120.github.io/assets/rblogger-criteria-img/`

The last thing to do is processing the output `rblogger-criteria.html` and replacing `<img src="../../assets` with `<img src="/assets` in the `<img>` tag. I do this with a simple bash script:

```bash
sed "s#<img src=\"../../assets#<img src=\"/assets#g" rblogger-criteria.html > temp
cat temp > rblogger-criteria.html
```

## A Demo

Below is a plot generated from an R Markdown code chunk, you can look at the path in `<img>` tag with the developer's tool or the [image source](/assets/rblogger-criteria-img/iris-1.png) to confirm the path of the image is just as mentioned above.
```{r iris}
library(ggplot2)
ggplot(iris, aes(Sepal.Length, Petal.Width)) +
    geom_point(aes(color = Species))
```


